Particle-in-cell simulation of ion and electron acceleration in trans-relativistic shocks.

### 2017-11-14 papers/Physica2017/intro.tex
paper

\section{Introduction}
One of the possible sources of high-energy cosmic rays are shock waves in astrophysical collisionless plasma due to the diffusive shock acceleration (DSA) mechanism, see \cite{Bell1978}, \cite{Blandford1978}. Strong magnetic fields are necessary for effective acceleration of charged particles and the observations of the synchrotron emission from supernova remnants (SNRs) show, that the magnetic field there is about 100 times stronger then the interstellar field (see for example \cite{Berezhko2003},\cite{Uchiyama2007}). So strong fields can be produced due to instabilities in anisotropic plasma. One of the most powerful methods to explore processes in collisionless plasma is the Particle-in-Cell simulation. In this work we explore instabilities in the plasma using this method.

>>>>

\section{Introduction}
One of the possible sources of high-energy cosmic rays are shock waves in astrophysical collisionless plasma due to the diffusive shock acceleration (DSA) mechanism, see \cite{Bell1978}, \cite{Blandford1978}. The particle acceleration on the shock front is sophisticated and nonlinear problem, because cosmic rays have strongly influence on the shock wave. The pressure of cosmic rays is comparable with pressure of background plasma and significatly modifies the front (see for example \cite{Bykov2014}). Also there are a instabilities in anisotropic plasma, which lead to strong field amplification and the observations of the synchrotron emission from supernova remnants (SNRs) show, that the magnetic field there is about 100 times stronger then the interstellar field (see for example \cite{Berezhko2003},\cite{Uchiyama2007}). One of the most powerful methods to explore processes in collisionless plasma is the Particle-in-Cell simulation.

### 2017-11-14 papers/Physica2017/shock.tex
paper

We developed the implicit particle-in-cell (PIC) code, presented in our previous paper \cite{Romansky2016}, based on the scheme suggested by Lapenta et al.~\cite{Lapenta2006} and improved for the relativistic case by Noguchi et al.\cite{Noguchi2007}.
Our code is fully three dimensional and parallelized with MPI technology, which is adapted for distributed computing and can be executed on a wide class of computers.

In simulation setup the homogeneous plasma flows in through the right boundary
and collides with the reflecting superconducting wall on the left boundary, causing the shock wave

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.86\textwidth]{fig/protons.eps} 
	\caption{Distribution of protons.}
	\label{protons}
\end{figure}
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.86\textwidth]{fig/electrons.eps} 
	\caption{Distribution of electrons.}
	\label{electrons}
\end{figure}

Simulations are one-dimensional and have following parameters: initial flow lorenz-factor $\gamma = 1.5$, number densities $n_e = 10^{-4} \rm{cm}^{-3}$, $n_p = 10^{-4} \rm{cm}^{-3}$ , temperature $5\cdot10^8 \rm{K}$, magnetic field $B = 10^{-4} \rm{G}$, the full size of the box $L = 2\cdot10^{12} \rm{cm}$, the number of cells $N=2\cdot10^4$. Electron mass is reduced to $m_e = \frac{m_p}{20}$. The full time of simulation is $T = 5000 {\omega_p}^{-1}$. $\theta$ is angle between flow velocity and magnetic field. We present results of particle spectrum in several simulations with different $\theta$.

One can see from Figures \ref{protons} and \ref{electrons} that spectrum of accelerated particles strongly depends on angle theta. If $\theta$ is less then critical value, defined by equation $c\cdot cos(\theta_{crit})=v_{shock}$, where all values are measured in upstream rest frame, particles can escape from the front and cross it several times to gain more energy. It is difficult to evaluate critical angle apriori, because we don't know compression relation of shock wave which will be created in simulation, but estimation for case of strong wave (comression relation $\sigma=4^{\circ}$) gets $\theta_{crit}=45$ in downstream frame. And it is consistent with results of our simulations.

>>>>

We developed the implicit particle-in-cell (PIC) code, based on the scheme suggested by Lapenta et al.~\cite{Lapenta2006} and improved for the relativistic case by Noguchi et al.\cite{Noguchi2007}.
Our code is fully three dimensional and parallelized with MPI technology, which is adapted for distributed computing and can be executed on a wide class of computers.

In simulation setup the homogeneous plasma, which flows in through the right boundary
and collides with the reflecting superconducting wall on the left boundary.

 Simulations are one-dimensional and have following parameters: initial flow lorenz-factor $\gamma = 1.5$, number densities $n_e = 10^{-4} \rm{cm}^{-3}$, $n_p = 10^{-4} \rm{cm}^{-3}$ , temperature $5\cdot10^8 \rm{K}$, magnetic field $B = 10^{-4} \rm{G}$, the full size of the box $L = 4\cdot10^{12} \rm{cm}$, the number of cells $N=4\cdot10^4$. Electron mass is reduced to $m_e = \frac{m_p}{20}$. The full time of simulation is $T = 5000 {\omega_p}^{-1}$. $\theta$ is angle between flow velocity and magnetic field. We present results of particle spectrum in several simulations with different $\theta$. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\textwidth]{fig/protons.eps} 
	\caption{Distribution of protons.}
	\label{protons}
\end{figure}
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\textwidth]{fig/electrons.eps} 
	\caption{Distribution of electrons.}
	\label{electrons}
\end{figure}

The results presented in Figures \ref{protons}-\ref{electrons_with_alpha} show that the values of $F(p)p^4$ for high energies are much greater for protons, than for electrons in both simulations. Also electrons need much more time to be accelerated and to form spectrum $\propto p^{-4}$. It means that protons are injected into the acceleration process more efficiently, and it is consistent with the work of Park et al. {\cite{Park2015}}. Also we have shown that minority of heavy ions increases the spectrum of protons and do not have influence on the spectrum of electrons. It can be explained by the fact, that heavy ions form the large scale turbulence and protons can efficiently scatter on this turbulence. Otherwise for electrons turbulence produced by protons is already enough large-scale  and the influence of heavy ions is neglectable.
